---
title: "Linear Regression"
author: Vishal Lala
output:
  html_document: 
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: haddock
    df_print: paged
  prettydoc::html_pretty:
    theme: cayman
    toc: true
    df_print: paged
  tufte::tufte_handout: default
  ioslides_presentation: default
  html_vignette: default
  pdf_document:
    toc: true
    toc_depth: 3
    highlight: haddock
  beamer_presentation: default
---

# Explore Data

## Structure

### Vector
Create a vector called income.
```{r}
set.seed(617)
income = round(runif(n = 10,min = 50000,max = 80000)/1000)*1000
income
#round(runif(n = 10,min = 50000,max = 80000)/1000)*10000
```

Here are a set of functions to explore income. 
class(), length(), max, which.max() 

Use above functions to answer the following questions about income

What is the class of income? 
```{r}
class(income)
length(income)
max(income)
which.max(income)
which.min(income)
```

How many elements does income contain?

```{r}
length(income)
```

What is the position of the maximum value in income?
```{r}
which.max(income)
```
which.max(income)

### Data Frame
Create a dataframe called df
```{r}
id = 1:10
set.seed(617)
age = round(runif(n=10,min = 25,max = 35))
set.seed(617)
gender = sample(c('Male','Female'),size=10,replace=T)
set.seed(617)
income = round(runif(n = 10,min = 50000,max = 80000)/1000)*1000
df = data.frame(id, age, gender, income)
```

Use the following functions to explore df
head(df)
nrow(df); ncol(df); dim(df)
names(df)
str(df)
View(df)

Display the first six elements in df.
```{r}
head(df)
tail(df)
nrow(df)
ncol(df)
dim(df)
names(df)
str(df)
```


List all the column names in df
```{r}
names(df)
```

Explore structure of df
```{r}
str(df)

```

## Subsetting
Examine the results from the following subsetting operations on income
income[2]
income[c(1,3)]
income[-c(1,3)]

```{r}
income[2]
income[c(1,3)]
income[-c(1,3)]
```

Examine the results from the following subsetting operations on df
df[1,1]
df[1,4]
df[c(1,2),3]
df[,4]
df[,c(1,4)]

```{r}

df[1,1]
df[1,4]
df[c(1,2),3]
df[,4]
df[,c(1,4)]



```

Fix the errors in the following subsetting operations
df(3)
```{r}
df[3]
```

income[3,4]
```{r}
income[c(3,4)]
```

df(4,4)
```{r}
df[4,4]

```

df[2,3,4]
```{r}
df[c(2,3),4]

```

df$income(2)
```{r}
df$income[2]
```

## Logical Tests

### Basic Tests
For an element, what do you expect from the following?
4 == 5
4 > 5
4 != 5

```{r}
4 == 5
4 > 5
4 != 5
```

Now, for a vector, what do you expect?
1:5 == 5
1:5 > 5
1:5 != 5
```{r}
1:5 == 5
1:5 > 5
which(1:5 == 5)
1:5 != 5
```

What do you expect from the following?
c(0, 1, 0, 1) == T
c('Vishal', 'Rohan', 'Nikhil') == 'Nikhil'

```{r}
c(0, 1, 0, 1) == T
c('Vishal', 'Rohan', 'Nikhil') == 'Nikhil'
```

### Sample Data
Create a set of vectors
```{r}
priceCoffee = c(3.49, 3.99,4.99,5.99)
x = 1:5
daysofweek = c('Monday','Tuesday','Wednesday','Thursday','Friday')
```

Now, use logical operators to answer the following questions
Is there a coffee priced at $3 or less?
```{r}
priceCoffee = c(3.49, 3.99,4.99,5.99)
as.logical(sum(priceCoffee < 3 ))
any(priceCoffee < 3)
min(priceCoffee) <= 3

```
Is there a coffee priced at exactly $3.49?
```{r}
as.logical(sum(priceCoffee == 3.49 ))
any(priceCoffee == 3.49)
```
Is x less than 4 and more than 2?
```{r}
#install.packages('ggplot2')
library(ggplot2)
data("diamonds")
str(diamonds)

```
Is ‘Sunday’ one of the daysofweek?
```{r}

```

### Diamonds Data
Load the diamonds dataset that ships with ggplot2
```{r}
library(ggplot2)
data(diamonds)
```
Now, use logical tests to answer the following.

How many diamonds are priced under $1000? (price is contained in the column price for the dataset diamonds)
```{r}
sum(diamonds$'price' < 1000)
table(diamonds$price<1000)

```

How many diamonds have x of 0? (x is contained in the column x for the dataset diamonds)
```{r}
table(diamonds$x==0)

```

Which diamonds have a price of $0? (price is contained in the column price for the dataset diamonds)
```{r}
which(diamonds$price == 0)
```

How many diamonds have a carat size greater than 4? (carat size is contained in the column carat in the dataset diamonds)
```{r}
sum(diamonds$carat >4 )
```

Does the dataset contain a diamond with a carat size of 7 or 8? (carat size is contained in the column carat in the dataset diamonds) [This is a tough one!]
```{r}
any(diamonds$carat > 7 )
any( diamonds$carat > 8 )

```

# Read Data
setwd('c:/myfiles/thatAwesomeClass/')
data = read.csv('data.csv')
```{r}

```

# Data Summarization
## Descriptive Measures

Examine measures of central tendency, and dispersion for diamonds. 
Measures of central tendency: mean(); median()
Measures of dispersion: range; max() - min(); sd(); var()

What is the average price of diamonds? 
```{r}
mean(diamonds$price)
```
What is the standard deviation for the price of a diamonds? 
```{r}

sd(diamonds$price)
```

What is the range of prices for diamonds?
```{r}
range(diamonds$price)
```


## Crosstabs
To create a table or crosstab of counts, use table()
For a crosstab that summarizes on some function, use tapply()

In the dataset, how many diamonds of each color are present? What about cut?
```{r}
table(diamonds$color)
```
What is the average price of diamonds of each color? 
```{r}
tapply(X = diamonds$price , INDEX =diamonds$color , FUN =  'mean')

```

Similarly, what is the average price of diamonds of each cut? 
```{r}

tapply(X = diamonds$price , INDEX =diamonds$cut , FUN =  'mean')
```

## Visual Summary
* Data  
* Aesthetic mapping (aes)  

  + Describes how variables are mapped onto graphical attributes  
  + Visual attribute of data including x-y axes, color, fill, shape, and alpha  
* Geometric objects (geom)  

  + Determines how values are rendered graphically, as bars (geom_bar), scatterplot (geom_point), line (geom_line), etc. 

Examine distribution of price by constructing a histogram
```{r}
library(ggplot2)
ggplot(data=diamonds, aes(x=price))+
  geom_histogram()
```

Distribution of price as a density curve
```{r}
ggplot(data=diamonds, aes(x=price))+
  geom_density()
```
Compare distribution of diamonds by cut.
```{r}
ggplot(data=diamonds, aes(x=price, color=cut))+
  geom_density(size=1.2)
```

Construct a bar chart to compare diamond price by cut?
```{r}
ggplot(data=diamonds, aes(x=cut, y=price, fill=cut))+
  geom_bar(position='dodge', stat='summary',fun.y='mean')+
  coord_flip()
```

Finally, examine relationship between price and carat using a scatterplot

```{r}
ggplot(data=diamonds, aes(x=carat,y=price))+
  geom_point(color='steelblue', size=0.8)

#install.packages('contrib.url')

```
